Robust-Adaptive Control of Linear Systems:\\ beyond Quadratic Costs

### 2020-06-01 main.tex
Fix typo
Draft new version of related work

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2020

% ready for submission
 \usepackage{neurips_2020}

@@

While there exists a vast body of literature on the control of uncertain systems, its majority is only concerned with the problem of stabilisation around a fixed reference state or trajectory, including $\cH_\infty$ control \citep[see, e.g.][]{Basar1996}, sliding-mode control \citep{Lu1997}, tube-MPC \citep{Limon2010} or system-level synthesis \citep{Dean2017,Dean2018}. Yet, we argue that many tasks cannot be framed as stabilisation problems (e.g. obstacle avoidance) and are better addressed in our setting of maximising an arbitrary utility function, where no prior result exists to our knowledge. Other safety frameworks exist, such as constraining the state $x$ in a given region $\mathbb{X}$ often chosen convex \citep[e.g.][]{Turchetta2016}, which again is not suitable for many tasks (e.g. these of \Cref{sec:experiments}). Therefore, we limit references to focus only on two lines of work closest to ours.

\paragraph{Linear systems with quadratic costs} Several approaches have been proposed for cumulative regret minimisation in the LQ problem. In the \emph{Optimism in the Face of Uncertainty} paradigm, the best possible dynamics within a high-confidence region is selected under a controllability constraint, to compute the corresponding optimal control in closed-form by solving a Riccati equation. \citep{abbasi-yadkori11a,Ibrahimi2013,Faradonbeh2017} show that this procedure achieves a $\tilde{\cO}\left(N^{1/2}\right)$ regret. Posterior sampling algorithms \citep{Ouyang2017,abeille18a} select candidate dynamics randomly instead, and obtain the same result. Other works use noise injection for exploration such as \citep{Dean2017,Dean2018}. However, neither optimism nor random exploration fit a critical setting, where ensuring safety requires instead to consider pessimistic outcomes. The work of \citet{Dean2017} is closest to our setting: after an offline estimation phase, they estimate a simple regret between a robust controller and the optimal performance. Our work mainly differs in that it addresses generic costs.

\paragraph{Robust Dynamic Programming with generic costs}
Robust optimisation with generic costs has been studied in the context of finite Markov Decision Processes with uncertain parameters by \citet{Iyengar2005}, \citet{Nilim2005} and \citet{Wiesemann2013}. They showed that the main results of Dynamic Programming can be extended to their robust counterparts only when the dynamics ambiguity set verifies a certain rectangularity property, related to time-varying uncertainty. Since our setting considers continuous states and time-invariant uncertainty, these methods do not apply.

@@

We abuse notations and define and include additional known terms in the measurement signal
$
%\begin{equation*}
%\label{eq:measurement}
    y(t) = \dot{x}(t) + C\nu(t) - A x(t) - Bu(t),
%\end{equation*}
$ 
to obtain a linear regression system
$
y_n = \Phi_n\theta + \eta_n.
$

>>>>

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2020
\PassOptionsToPackage{numbers, compress}{natbib}
% ready for submission
 \usepackage{neurips_2020}

@@

The control of uncertain systems is a long-standing problem, to which a vast body of literature is dedicated. Existing work is mostly concerned with the problem of \emph{stabilisation} around a fixed reference state or trajectory, including approaches such as $\cH_\infty$ control \citep[see, e.g.][]{Basar1996}, sliding-mode control \citep{Lu1997} or system-level synthesis \citep{Dean2017,Dean2018}. This paper fits in the popular MPC framework, for which adaptive data-driven schemes have been developed to deal with model uncertainty \citep{Sastry1990,Tanaskovic2014}. 
Need for guarantees. a common safety objective is the \emph{robust constraint satisfaction}, where the state $x$ is constrained into a region $\mathbb{X}$, often chosen convex \citep[e.g.][]{Turchetta2016,Lorenzen2017,Kohler2019,Lu2019} a line of work commonly referred to as Tube-MPC algorithms.

To go beyond stabilization, we consider is the minimax objective. 
Yet, we argue that many tasks cannot be framed as stabilisation problems (e.g. obstacle avoidance) and are better addressed in our setting of maximising an utility function. 
It has been mostly studied in two particular cases: (1) linear systems with quadratic costs, and (2) finite state space. 
Another kind of guarantee, is the minimax

Finite states
Robust optimisation with generic costs has been studied in the context of finite Markov Decision Processes with uncertain parameters by \citet{Iyengar2005}, \citet{Nilim2005} and \citet{Wiesemann2013}. They showed that the main results of Dynamic Programming can be extended to their robust counterparts only when the dynamics ambiguity set verifies a certain rectangularity property. Since we consider continuous states, these methods do not apply.

Linear dynamics + Quadratic costs
This has been vastly studied for quadratic costs, where several approaches have been proposed for cumulative regret minimisation in the LQ problem. In the \emph{Optimism in the Face of Uncertainty} paradigm, the best possible dynamics within a high-confidence region is selected under a controllability constraint, to compute the corresponding optimal control in closed-form by solving a Riccati equation. \citep{abbasi-yadkori11a,Ibrahimi2013,Faradonbeh2017} show that this procedure achieves a $\tilde{\cO}\left(N^{1/2}\right)$ regret. Posterior sampling algorithms \citep{Ouyang2017,abeille18a} select candidate dynamics randomly instead, and obtain the same result. Other works use noise injection for exploration such as \citep{Dean2017,Dean2018}. However, neither optimism nor random exploration fit a critical setting, where ensuring safety requires instead to consider pessimistic outcomes. The work of \citet{Dean2017} is closest to our setting: after an offline estimation phase, they estimate a simple regret between a robust controller and the optimal performance. 
Our work differs in that it addresses arbitrary bounded costs, rather than quadratic costs only.

%While there exists a vast body of literature on the control of uncertain systems, its majority is only concerned with the problem of stabilisation around a fixed reference state or trajectory, including $\cH_\infty$ control \citep[see, e.g.][]{Basar1996}, sliding-mode control \citep{Lu1997}, tube-MPC \citep{Limon2010} or system-level synthesis \citep{Dean2017,Dean2018}. Yet, we argue that many tasks cannot be framed as stabilisation problems (e.g. obstacle avoidance) and are better addressed in our setting of maximising an arbitrary utility function, where no prior result exists to our knowledge. Other safety frameworks exist, such as constraining the state $x$ in a given region $\mathbb{X}$ often chosen convex \citep[e.g.][]{Turchetta2016}, which again is not suitable for many tasks (e.g. these of \Cref{sec:experiments}). Therefore, we limit references to focus only on two lines of work closest to ours.
%
%\paragraph{Linear systems with quadratic costs} Several approaches have been proposed for cumulative regret minimisation in the LQ problem. In the \emph{Optimism in the Face of Uncertainty} paradigm, the best possible dynamics within a high-confidence region is selected under a controllability constraint, to compute the corresponding optimal control in closed-form by solving a Riccati equation. \citep{abbasi-yadkori11a,Ibrahimi2013,Faradonbeh2017} show that this procedure achieves a $\tilde{\cO}\left(N^{1/2}\right)$ regret. Posterior sampling algorithms \citep{Ouyang2017,abeille18a} select candidate dynamics randomly instead, and obtain the same result. Other works use noise injection for exploration such as \citep{Dean2017,Dean2018}. However, neither optimism nor random exploration fit a critical setting, where ensuring safety requires instead to consider pessimistic outcomes. The work of \citet{Dean2017} is closest to our setting: after an offline estimation phase, they estimate a simple regret between a robust controller and the optimal performance. Our work mainly differs in that it addresses generic costs.
%
%\paragraph{Robust Dynamic Programming with generic costs}
%Robust optimisation with generic costs has been studied in the context of finite Markov Decision Processes with uncertain parameters by \citet{Iyengar2005}, \citet{Nilim2005} and \citet{Wiesemann2013}. They showed that the main results of Dynamic Programming can be extended to their robust counterparts only when the dynamics ambiguity set verifies a certain rectangularity property, related to time-varying uncertainty. Since our setting considers continuous states and time-invariant uncertainty, these methods do not apply.

@@

We slightly abuse notations and include additional known terms in the measurement signal
$
%\begin{equation*}
%\label{eq:measurement}
    y(t) = \dot{x}(t) + C\nu(t) - A x(t) - Bu(t),
%\end{equation*}
$ 
to obtain a linear regression system
$
y_n = \Phi_n\theta + \eta_n.
$
